---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

-   Tools \> Global Options \> Code \> Insert matching parens/quotes

-   Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### Load Library

```{r}
library(tidyverse)
library(ggplot2)

```

### Some Scripting

```{r}
quiz1 <- 88

quiz2 <- 94

quiz3 <- 79

quiz1 

mean(quiz1 + quiz2 + quiz3) #adds everything up
mean(quiz1, quiz2, quiz3) #only returns the first arg

mean(c(quiz1, quiz2, quiz3)) # CORRECT WAY, c() means combine or concatenate

list <- c(quiz1, quiz2, quiz3)

mean(list)

average <- mean(list)

quiz3 <- 99
quiz3

average

list <- c(quiz1, quiz2, quiz3)
average <- mean(list)
average
round(average, 2)
```

### More Scores!

```{r}

scores <- c(88, 94,79, 91,85)
scores

mean(scores)
min(scores)
max(scores)
length(scores)

summary(scores)
### remember summary scores 

students <- c("a", "b", "c", "d", "e")
students

#bar plot
barplot(scores, names.arg = students, ylim = c(0,100), main = "Title", xlab = "pupils", y = "what they got")
```

## Introducing `ggplot2`

#### Make a data frame

```{r}

grades_df <- data.frame(student = students, score = scores)

grades_df

summary(grades_df)

nrow(grades_df) #number of rows

ncol(grades_df) #number of columns

dim(grades_df) #number of data frame dimension


```

#### Try ggplot

```{r}
ggplot(data = grades_df, aes(x = student, y = score)) #if run alone this is foundation

ggplot(data = grades_df, aes(x = reorder(student, -score), y = score, group = 1)) +
  geom_col(fill = "steelblue") + 
  geom_point() + 
  geom_line() +
  theme_minimal() + 
  labs(title = "Best Graph Ever", 
       subtitle = "some quiz scores",
       x = "da students",
       y = "what they scored on their quiz")

# group in the aes() connects the line geom_line()
#reorder(after the x puts it in descending order)

```

#### Mutate the Dataframe

```{r}
library(tidyverse)
grades_df <- grades_df %>% 
  mutate(study_hours = c(2,5,1,4,3))


#pipe operator (command-shift-m)
```

#### Make a scatterplot

```{r}
ggplot(data = grades_df, aes(x = study_hours, y = scores)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() +
  labs(title = "The Impact of Studying on Test Scores", subtitle = "by Kyle Wong", x = "Time Spent Studying", y = "Test Scores")

##dont use fill for lines (use color instead), dont use reorder 


```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}
1 + 1 + 1

quiz1 + quiz2

1 + 1 + 1 - 3

0.1 + 0.1 + 0.1 - 0.3

options(scipen = 999)

print(0.1, 20)
print(0.3, 20)
print(0.5, 20)


```

What did you observe?

> Answer:
>
> I expected to see 0 for the 0.1 + 0.1 + 0.1 - 0.3 expression.

Why do you think this is happening? (You can look online for help)

> Answer:

We typically calculate and process numbers through base-10, which produces cleaner numbers. R language processes in base-2, or binary(0s and 1s) which leads to rounding errors.
